---
title: "Solving a parabolic equation"
date: "Created: 20-04-2025. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```


Let us set some global options for all code chunks in this document.


```{r}
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
```




```{r}
# inla.upgrade(testing = TRUE)
# remotes::install_github("inlabru-org/inlabru", ref = "devel")
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")
library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)
library(grateful)

library(plotly)
```


We want to solve the fractional diffusion equation
\begin{equation}
\label{eq:maineq}
    \partial_t u+(\kappa^2-\Delta_\Gamma)^{\frac{\alpha}{2}} u=f \text { on } \Gamma \times(0, T), \quad u(0)=u_0 \text { on } \Gamma,
\end{equation}
where $u$ satisfies the Kirchhoff vertex conditions
\begin{equation}
\label{eq:Kcond}
    \left\{\phi\in C(\Gamma)\;\Big|\; \forall v\in V: \sum_{e\in\mathcal{E}_v}\partial_e \phi(v)=0 \right\}
\end{equation}

Let us start by building a graph.

```{r}
graph <- metric_graph$new(perform_merges = TRUE, 
                          tolerance = list(edge_edge = 1e-3, 
                                           vertex_vertex = 1e-3, 
                                           edge_vertex = 1e-3))
graph$plot()
graph$build_mesh(h = 0.05)
```

Let $\alpha\in(0,2]$ and $U_h^\tau$ denote the sequence of approximations of the solution to the weak form of problem \eqref{eq:maineq} at each time step on a mesh indexed by $h$. Let $z=0$ and $U^0_h = P_hu_0$. For $k=0,\dots, N-1$, $U_h^{k+1}\in V_h$ solves the following scheme
\begin{align}
\label{system:fully_discrete_scheme}
        \langle\delta U_h^{k+1},\phi\rangle + \mathfrak{a}(U_h^{k+1},\phi) = \langle f^{k+1},\phi\rangle ,\quad\forall\phi\in V_h,
\end{align}
where $f^{k+1} = \displaystyle\dfrac{1}{\tau}\int_{t_k}^{t^{k+1}}f(t)dt$.
The solution can be represented as 
\begin{align*}
    U_h^k(s) =  \sum_{j=1}^{N_h}u_j^k\psi_j(s)
\end{align*}
Replacing this into \eqref{system:fully_discrete_scheme} yields the following linear system
\begin{align*}
    \sum_{j=1}^{N_h}u_j^{k+1}[(\psi_j,\psi_i)_{L_2(\Gamma)}+ \tau\mathfrak{a}(\psi_j,\psi_i)] = \sum_{j=1}^{N_h}u_j^{k}(\psi_j,\psi_i)_{L_2(\Gamma)}+\tau( f^{k+1},\psi_i)_{L_2(\Gamma)}
\end{align*}
for $i = 1,\dots, N_h$. In matrix notation,
\begin{align}
\label{diff_eq_discrete}
    (C+\tau A)U^{k+1} = CU^k+\tau F^{k+1},
\end{align}
where $C$ has entries $C_{ij} = (\psi_j,\psi_i)_{L_2(\Gamma)}$, $A$ has entries $A_{ij} = \mathfrak{a}(\psi_j,\psi_i)$, $U^k$ has entries $u_j^k$, and $F^k$ has entries $( f^{k},\psi_i)_{L_2(\Gamma)}$. By multiplying both sides by $A^{-1}$ and considering its operator-based rational approximation $P_\ell^{-1}P_r$, we arrive at $(P_rC+\tau P_\ell)U^{k+1} = P_r(CU^k+\tau F^{k+1})$.

```{r}
# Compute the FEM matrices
graph$compute_fem()
G <- graph$mesh$G
C <- graph$mesh$C
x <- graph$mesh$V[, 1]
y <- graph$mesh$V[, 2]
# Initial condition
U_0 <- 10*exp(-((x-4)^2 + (y-4)^2))
# Define the time step
time_step <- 0.1
# Define the right-hand side function
fun <- function(t) {return(sin(t)*((x-4)^2 - (y-4)^2))}
# Define the time discretization
time_seq <- seq(0,pi, by = time_step)
# Compute the right-hand side function at each time step
fun_mat <- do.call(cbind, lapply(time_seq, fun))
# Define the parameters
kappa <- 1
L <- kappa^2*C + G
alpha <- 0.8
beta <- alpha/2
op <- fractional.operators(L, beta, C, scale.factor = kappa^2, m = 1)
Pl <- op$Pl
Pr <- op$Pr
funF <- C %*% fun_mat 
```


```{r}
# Precompute the LHS matrix
LHS <- Pr %*% C + time_step * Pl
# Initialize U matrix to store solution at each time step
U_mat <- matrix(NA, nrow = nrow(C), ncol = length(time_seq))
U_mat[, 1] <- U_0

# Time-stepping loop
for (k in 1:(length(time_seq) - 1)) {
  RHS <- Pr %*% (C %*% U_mat[, k] + time_step * funF[, k + 1])
  U_mat[, k + 1] <- as.matrix(solve(LHS, RHS))
}
```


```{r}
# Plot the initial condition
p_ini <- graph$plot_function(X = U_0, 
                             vertex_size = 1, 
                             type = "plotly", 
                             edge_color = "black", 
                             edge_width = 3, 
                             line_color = "blue", 
                             line_width = 3)
p_ini
# Plot the movie of f
p_f <- graph$plot_movie(fun_mat)
p_f$x$layout$scene$xaxis$range <- range(x)
p_f$x$layout$scene$yaxis$range <- range(y)
p_f$x$layout$scene$zaxis$range <- range(fun_mat)
p_f
# Plot the movie of the solution
p_sol <- graph$plot_movie(U_mat)
p_sol$x$layout$scene$xaxis$range <- range(x)
p_sol$x$layout$scene$yaxis$range <- range(y)
p_sol$x$layout$scene$zaxis$range <- range(U_mat)
p_sol
```



# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
